---
title: "Data Management Cheat Sheet"
author: "David D. Liebowitz"
date: "EDUC 641: Fall 2022"
site: distill::distill_website
output: 
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setting up project and directory structure

Open your RStudio, create a project and save it. Go to the root directory of the project and create folders named: "Code", "Data", "Figures" and "Tables." Download the [life_expectancy.csv dataset](./data/life_expectancy.csv) and store it in the folder "Data". Create an R script (or .Rmd) file in the Code folder. 

## Load relevant packages
```{r, echo=T, results="hide"}
# Call one package
library(pacman)
# Call multiple packages at once
p_load(tidyverse, here, modelsummary)
```

## Read the data in
```{r, echo=T}
df <- read.csv(here("data/life_expectancy.csv"))
# There are other packages to read in data that are much faster and more flexible
```

```{r, echo=F}
df <- df %>%
  janitor::clean_names() %>%
  mutate(status = case_when(status == "Developed" ~ 1,
                            status == "Developing" ~ 0))
```

## Extract cases
```{r, echo=T}
df <- filter(df, year == 2015)
```

## Extract variables
```{r, echo=T}
df <- df %>% select(country, status, schooling, life_expectancy)
```

## Rename variable
```{r, echo=T}
df <- rename(df, region = country)
```

## Create a new variable based on existing one(s)
```{r, echo=T}
# Replace existing variable
df <- df %>%  
      mutate(life_expectancy = round(life_expectancy, digits = 0))
# Create a new one
df <- df %>%
      mutate(life2 = life_expectancy * life_expectancy)
```

## Inspect your data
```{r, echo=T}
head(df)
str(df)
```

## Transform variables, as needed
```{r, echo=T}
df$status <- factor(df$status,
                   levels = c(0, 1),
                   labels = c("Developing", "Developed"))
```


## Examine missingness
```{r, echo=T}
# One at a time
sum(is.na(df$schooling))

# All at once
sapply(df, function(x) sum(is.na(x)))
```

## Decide on listwise/pairwise deletion
```{r, echo=T}
# In this case, we'll use listwise
df <- filter(df, !is.na(schooling))
```

## Calculate summary statistics
```{r, echo=T}
mean(df$life_expectancy)
median(df$life_expectancy)
sd(df$life_expectancy)
summary(df$life_expectancy)
table(df$status, exclude=NULL)
```

## Visualize variables (categorical)
```{r, echo=T}
counts <- table(df$status)
barplot(counts)
```

## Visualize variables (continuous)
```{r, echo=T}
hist(df$life_expectancy)
plot(density(df$schooling))
```

## Create descriptive table (continuous)
```{r, echo=T}
datasummary_skim(df)
```

## Create descriptive table (categorical)
```{r, echo=T}
datasummary_skim(df, type="categorical")
```